Monster or Lambda?  A likelihood analysis

### 2023-04-23 inst/f1points.Rmd
work on rebuttal (constency wta vs top1), issue #197

```{r, label=calcpoints_real_2017, echo=FALSE, cache=TRUE}
dp2017 <- ordertable2points(as.ordertable(F1_table_2017[seq_len(top),]),p)
dp2017 <- sort(dp2017,decreasing=TRUE)
dp2017[] <- seq_along(dp2017)
m2017 <- maxp(H2017)
m2017 <- sort(m2017,decreasing=TRUE)
m2017[] <- seq_along(m2017)

@@

@@

```{r,label=definemultiplemontecarlo, echo=FALSE}
load("formula1_results_2016.rda") # this created by running f1points_Omaker.R; file contains R variables OO and pointslist
OO2016 <- OO

load("formula1_results_2017.rda") # this created by running f1points_Omaker.R; file contains R variables OO and pointslist
OO2017 <- OO

load("formula1_results_2018.rda") # this created by running f1points_Omaker.R; file contains R variables OO and pointslist
OO2018 <- OO

load("formula1_results_2019.rda") # this created by running f1points_Omaker.R; file contains R variables OO and pointslist
OO2019 <- OO
```

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='mean correct places',axes=FALSE)
axis(1,at=1:6,labels=names(formula1_points_systems(top)[wanted]))
axis(2,at=1:4)
points(s2016[wanted,1],pch=1,type='b',col='red')
points(s2017[wanted,1],pch=2,type='b',col='orange')
points(s2018[wanted,1],pch=3,type='b',col='green')
points(s2019[wanted,1],pch=4,type='b',col='blue')
#legend("topright",lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=6,y=s2016[wanted[6],1],pos=4,"2016",col='red')
text(x=6,y=s2017[wanted[6],1],pos=4,"2017",col='orange')
text(x=6,y=s2018[wanted[6],1],pos=4,"2018",col='green')
text(x=6,y=s2019[wanted[6],1],pos=4,"2019",col='blue')
```

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='probability of winner correct',axes=FALSE)
axis(1,at=1:6,labels=names(formula1_points_systems(top)[wanted]))
axis(2)
points(s2016[wanted,2],pch=1,type='b',col='red')
points(s2017[wanted,2],pch=2,type='b',col='orange')
points(s2018[wanted,2],pch=3,type='b',col='green')
points(s2019[wanted,2],pch=4,type='b',col='blue')
#legend(x=3.4,y=0.67,lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=6,y=s2016[wanted[6],2],pos=4,"2016",col='red')
text(x=6,y=s2017[wanted[6],2],pos=4,"2017",col='orange')
text(x=6,y=s2018[wanted[6],2],pos=4,"2018",col='green')
text(x=6,y=s2019[wanted[6],2],pos=4,"2019",col='blue')
```

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='probability of total rank order correct',axes=FALSE)
axis(1,at=1:6,labels=names(formula1_points_systems(top)[wanted]))
axis(2)
points(s2016[wanted,3],pch=1,type='b',col='red')
points(s2017[wanted,3],pch=2,type='b',col='orange')
points(s2018[wanted,3],pch=3,type='b',col='green')
points(s2019[wanted,3],pch=4,type='b',col='blue')
#legend("topleft",lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=5,y=s2016[wanted[5],3],pos=3,"2016",col='red')
text(x=5,y=s2017[wanted[5],3],pos=1,"2017",col='orange')
text(x=5,y=s2018[wanted[5],3],pos=3,"2018",col='green')
text(x=5,y=s2019[wanted[5],3],pos=3,"2019",col='blue')
```

>>>>

```{r, label=calcpoints_real_2017, echo=FALSE, cache=TRUE}
dp2017 <- ordertable2points(as.ordertable(F1_table_2017[seq_len(top),]),p)
dp2017 <- sort(dp2017,decreasing=TRUE)
dp2017[] <- seq_along(dp2017)
m2017 <- maxp(H2017)
m2017 <- sort(m2017,decreasing=TRUE)
m2017[] <- seq_along(m2017)
```

@@

```{r label=zipf_order_vs_likelihood_order_2016,cache=TRUE}
par(pty='s')        # square plot
ox <- dp2016z
oy <- ordertrans(m2016,names(dp2016z))
plot(ox,oy,asp=1,pty='s',xlim=c(0,11.2),ylim=c(0,11.2),pch=16,
    xlab="points order (Zipf)",ylab="likelihood order",main='2016 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray',cex=0.6) }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)
```

```{r label=zipf_order_vs_likelihood_order_2017,cache=TRUE}
par(pty='s')        # square plot
ox <- dp2017z
oy <- ordertrans(m2017,names(dp2017z))
plot(ox,oy,asp=1,pty='s',xlim=c(0,11.2),ylim=c(0,11.2),pch=16,
    xlab="points order (Zipf)",ylab="likelihood order",main='2017 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray', cex=0.6) }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)
```

```{r label=zipf_order_vs_likelihood_order_2018,cache=TRUE}
par(pty='s')        # square plot
ox <- dp2018z
oy <- ordertrans(m2018,names(dp2018z))
plot(ox,oy,asp=1,pty='s',xlim=c(0,11.2),ylim=c(0,11.2),pch=16,
    xlab="points order (Zipf)",ylab="likelihood order",main='2018 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray', cex=0.6) }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)
```

```{r label=zipf_order_vs_likelihood_order_2019,cache=TRUE}
par(pty='s')        # square plot
ox <- dp2019z
oy <- ordertrans(m2019,names(dp2019z))
plot(ox,oy,asp=1,pty='s',xlim=c(0,11.2),ylim=c(0,11.2),pch=16,
    xlab="points order (Zipf)",ylab="likelihood order",main='2019 season')
par(xpd=TRUE)       # allow drivers' names to appear outside plotting region
for(i in seq_along(ox)){text(ox[i],oy[i],names(ox)[i],pos=4,col='gray',cex=0.6) }
par(xpd=FALSE)      # stop diagonal line from protruding beyond plotting region
abline(0,1)
```

@@

```{r,label=definemultiplemontecarlo, echo=FALSE}
load("formula1_results_2016.rda") # this created by running f1points_Omaker.R; file contains R variables OO and pointslist
rownames(OO)[3] <- "wta"
OO2016 <- OO

load("formula1_results_2017.rda") # this created by running f1points_Omaker.R; file contains R variables OO and pointslist
rownames(OO)[3] <- "wta"
OO2017 <- OO

load("formula1_results_2018.rda") # this created by running f1points_Omaker.R; file contains R variables OO and pointslist
rownames(OO)[3] <- "wta"
OO2018 <- OO

load("formula1_results_2019.rda") # this created by running f1points_Omaker.R; file contains R variables OO and pointslist
rownames(OO)[3] <- "wta"
OO2019 <- OO
```

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='mean correct places',axes=FALSE)
jjlabels <- names(formula1_points_systems(top)[wanted])
jjlabels[5] <- "wta"
axis(1,at=1:6,labels=jjlabels)
axis(2,at=1:4)
points(s2016[wanted,1],pch=1,type='b',col='red')
points(s2017[wanted,1],pch=2,type='b',col='orange')
points(s2018[wanted,1],pch=3,type='b',col='green')
points(s2019[wanted,1],pch=4,type='b',col='blue')
#legend("topright",lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=6,y=s2016[wanted[6],1],pos=4,"2016",col='red')
text(x=6,y=s2017[wanted[6],1],pos=4,"2017",col='orange')
text(x=6,y=s2018[wanted[6],1],pos=4,"2018",col='green')
text(x=6,y=s2019[wanted[6],1],pos=4,"2019",col='blue')
```

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='probability of winner correct',axes=FALSE)
axis(1,at=1:6,labels=jjlabels)
axis(2)
points(s2016[wanted,2],pch=1,type='b',col='red')
points(s2017[wanted,2],pch=2,type='b',col='orange')
points(s2018[wanted,2],pch=3,type='b',col='green')
points(s2019[wanted,2],pch=4,type='b',col='blue')
#legend(x=3.4,y=0.67,lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=6,y=s2016[wanted[6],2],pos=4,"2016",col='red')
text(x=6,y=s2017[wanted[6],2],pos=4,"2017",col='orange')
text(x=6,y=s2018[wanted[6],2],pos=4,"2018",col='green')
text(x=6,y=s2019[wanted[6],2],pos=4,"2019",col='blue')
```

@@

matplot(jj[wanted,],type="n",xlab='points system',ylab='probability of total rank order correct',axes=FALSE)
axis(1,at=1:6,labels=jjlabels)
axis(2)
points(s2016[wanted,3],pch=1,type='b',col='red')
points(s2017[wanted,3],pch=2,type='b',col='orange')
points(s2018[wanted,3],pch=3,type='b',col='green')
points(s2019[wanted,3],pch=4,type='b',col='blue')
#legend("topleft",lty=0,pch=1:4,legend=c("2016","2017","2018","2019"))
par(xpd=TRUE)
text(x=5,y=s2016[wanted[5],3],pos=3,"2016",col='red')
text(x=5,y=s2017[wanted[5],3],pos=1,"2017",col='orange')
text(x=5,y=s2018[wanted[5],3],pos=3,"2018",col='green')
text(x=5,y=s2019[wanted[5],3],pos=3,"2019",col='blue')
```
